A Logical Solution of the Linearity Challenge

### 2025-12-02 paper/Reduction.lagda.tex
restored reductions

\begin{code}[hide]
open import Data.Bool using (Bool; true; false; if_then_else_)
open import Data.Product using (_×_; _,_; ∃; ∃-syntax)
open import Relation.Binary.PropositionalEquality using (_≡_; refl; sym; cong; cong₂)
open import Data.List.Base using (List; []; _∷_; [_]; _++_)
open import Data.List.Properties using (++-assoc)

@@

\begin{code}[hide]
weakening : ∀{Γ Γ₁ Γ₂} (un : Un Γ₁) → Γ ≃ Γ₁ + Γ₂ → Process Γ₂ → Process Γ
weakening un p P = #process (+++# p) (aux un P)
  where
    aux : ∀{Γ₁ Γ₂} (un : Un Γ₁) → Process Γ₂ → Process (Γ₁ ++ Γ₂)
    aux un-[] P = P
    aux (un-∷ un) P = weaken (split-l +-unit-l) (aux un P)

contraction : ∀{Γ Γ₁ Γ₂} (un : Un Γ₁) → Γ ≃ Γ₁ + Γ₂ → Process (Γ₁ ++ Γ) → Process Γ
contraction un p P = #process (+++# p) (aux un (#process (#left (#sym (+++# p))) P))
  where
    aux : ∀{Γ₁ Γ₂} → Un Γ₁ → Process (Γ₁ ++ Γ₁ ++ Γ₂) → Process (Γ₁ ++ Γ₂)
    aux un-[] P = P
    aux {¿ A ∷ Γ₁} {Γ₂} (un-∷ un) P with contract (split-l +-unit-l) (#process (#shift {¿ A ∷ Γ₁} {¿ A} {Γ₁ ++ Γ₂}) P)
    ... | P₁ rewrite sym (++-assoc (¿ A ∷ Γ₁) Γ₁ Γ₂) with #process (#sym (#shift {Γ₁ ++ Γ₁} {¿ A})) P₁
    ... | P₂ rewrite ++-assoc Γ₁ Γ₁ (¿ A ∷ Γ₂) with aux un P₂
    ... | P₃ = #process #shift P₃
\end{code}

@@

\begin{code}
  r-link      : ∀{Δ A B} {P : Process (B ∷ Δ)}
                (d : Dual A B) (e : Dual A B) (p : Γ ≃ B , Δ) →
                cut d p (link e (split-l (split-r split-e))) P ↝ #process (#cons p) P
\end{code}

@@

\begin{code}
  r-close     : ∀{P : Process Γ} (p₀ : Γ ≃ [] + Γ) (q₀ : Γ ≃ [] + Γ) →
                cut d-𝟙-⊥ p₀ close (wait (split-l q₀) P) ↝ P
\end{code}

@@

\begin{code}
  r-select-l  : ∀{Γ₁ Γ₂ A A′ B B′}
                {P : Process (A ∷ Γ₁)} {Q : Process (A′ ∷ Γ₂)} {R : Process (B′ ∷ Γ₂)}
                (d : Dual A A′) (e : Dual B B′)
                (p : Γ ≃ Γ₁ + Γ₂) (p₀ : Γ₁ ≃ [] + Γ₁) (q₀ : Γ₂ ≃ [] + Γ₂) →
                cut (d-⊕-& d e) p
                    (select true (split-l p₀) P)
                    (case (split-l q₀) Q R) ↝ cut d p P Q
\end{code}

@@

\begin{code}[hide]
  r-select-r  :
    ∀{Γ₁ Γ₂ A A′ B B′}
    {P : Process (B ∷ Γ₁)} {Q : Process (A′ ∷ Γ₂)} {R : Process (B′ ∷ Γ₂)}
    (d : Dual A A′) (e : Dual B B′) (p : Γ ≃ Γ₁ + Γ₂) (p₀ : Γ₁ ≃ [] + Γ₁) (q₀ : Γ₂ ≃ [] + Γ₂) →
    cut (d-⊕-& d e) p
        (select false (split-l p₀) P)
        (case (split-l q₀) Q R) ↝ cut e p P R

  r-fork      :
    ∀{Γ₁ Γ₂ Γ₃ Δ A B A′ B′}
    {P : Process (A ∷ Γ₁)} {Q : Process (B ∷ Γ₂)} {R : Process (B′ ∷ A′ ∷ Γ₃)}
    (d : Dual A A′) (e : Dual B B′) (p : Γ ≃ Δ + Γ₃) (p₀ : Γ₃ ≃ [] + Γ₃)
    (q : Δ ≃ Γ₁ + Γ₂) (q₀ : Δ ≃ [] + Δ) →
    let _ , p′ , q′ = +-assoc-l p q in
    cut (d-⊗-⅋ d e) p
        (fork (split-l q₀) q P Q)
        (join (split-l p₀) R) ↝ cut d q′ P (cut e (split-r p′) Q R)

  r-client    :
    ∀{Γ₁ Γ₂ A A′}
    {P : Process (A ∷ Γ₁)} {Q : Process (A′ ∷ Γ₂)} (d : Dual A A′)
    (p : Γ ≃ Γ₁ + Γ₂) (p₀ : Γ₁ ≃ [] + Γ₁) (q₀ : Γ₂ ≃ [] + Γ₂) (un : Un Γ₁) →
    cut (d-!-? d) p
      (server (split-l p₀) un P)
      (client (split-l q₀) Q) ↝ cut d p P Q

  r-weaken    :
    ∀{Γ₁ Γ₂ A A′}
    {P : Process (A ∷ Γ₁)} {Q : Process Γ₂}
    (d : Dual A A′) (p : Γ ≃ Γ₁ + Γ₂) (p₀ : Γ₁ ≃ [] + Γ₁) (q₀ : Γ₂ ≃ [] + Γ₂) (un : Un Γ₁) →
    cut (d-!-? d) p
        (server (split-l p₀) un P)
        (weaken (split-l q₀) Q) ↝ weakening un p Q

  r-contract  :
    ∀{Γ₁ Γ₂ A A′}
    {P : Process (A ∷ Γ₁)} {Q : Process (¿ A′ ∷ ¿ A′ ∷ Γ₂)}
    (d : Dual A A′) (p : Γ ≃ Γ₁ + Γ₂) (p₀ : Γ₁ ≃ [] + Γ₁) (q₀ : Γ₂ ≃ [] + Γ₂) (un : Un Γ₁) →
    cut (d-!-? d) p
      (server (split-l p₀) un P)
      (contract (split-l q₀) Q) ↝
      contraction un p
        (cut (d-!-? d) ++≃+
             (server (split-l p₀) un P)
             (cut (d-!-? d) (split-r p) (server (split-l p₀) un P) Q))
\end{code}

@@

\begin{code}
  r-cut       : ∀{Γ₁ Γ₂ A B} {P Q : Process (A ∷ Γ₁)} {R : Process (B ∷ Γ₂)}
                (d : Dual A B) (q : Γ ≃ Γ₁ + Γ₂) → P ↝ Q →
                cut d q P R ↝ cut d q Q R

  r-cong      : ∀{P R Q : Process Γ} → P ⊒ R → R ↝ Q → P ↝ Q
\end{code}
\end{AgdaAlign}

>>>>

\begin{code}[hide]
{-# OPTIONS --rewriting #-}
open import Data.Bool using (Bool; true; false; if_then_else_)
open import Data.Product using (_×_; _,_; ∃; Σ; Σ-syntax; ∃-syntax)
open import Relation.Binary.PropositionalEquality using (_≡_; refl; sym; cong; cong₂)
open import Data.List.Base using (List; []; _∷_; [_]; _++_)
open import Data.List.Properties using (++-assoc)

@@

\begin{code}[hide]
weakening : ∀{Γ Γ₁ Γ₂} (un : Un Γ₁) → Γ ≃ Γ₁ + Γ₂ → Process Γ₂ → Process Γ
weakening un p P = ↭process (↭concat p) (aux un P)
  where
    aux : ∀{Γ₁ Γ₂} (un : Un Γ₁) → Process Γ₂ → Process (Γ₁ ++ Γ₂)
    aux un-[] P = P
    aux (un-∷ un) P = weaken (< ≫) (aux un P)

contraction : ∀{Γ Γ₁ Γ₂} (un : Un Γ₁) → Γ ≃ Γ₁ + Γ₂ → Process (Γ₁ ++ Γ) → Process Γ
contraction un p P = ↭process (↭concat p) (aux un (↭process (↭left (↭sym (↭concat p))) P))
  where
    aux : ∀{Γ₁ Γ₂} → Un Γ₁ → Process (Γ₁ ++ Γ₁ ++ Γ₂) → Process (Γ₁ ++ Γ₂)
    aux un-[] P = P
    aux {`? A ∷ Γ₁} {Γ₂} (un-∷ un) P with contract (< ≫) (↭process (↭shift {`? A} {`? A ∷ Γ₁} {Γ₁ ++ Γ₂}) P)
    ... | P₁ rewrite sym (++-assoc (`? A ∷ Γ₁) Γ₁ Γ₂) with ↭process (↭sym (↭shift {`? A} {Γ₁ ++ Γ₁})) P₁
    ... | P₂ rewrite ++-assoc Γ₁ Γ₁ (`? A ∷ Γ₂) with aux un P₂
    ... | P₃ = ↭process ↭shift P₃
\end{code}

@@

\begin{code}
  r-link      : ∀{Δ A P} (p : Γ ∋ dual A ⊳ Δ) →
                cut {A} p (link (< > •)) P ↝ ↭process (↭concat p) P
\end{code}

@@

\begin{code}
  r-close     : ∀{P} (p₀ : Γ ≃ [] + Γ) (q₀ : Γ ≃ [] + Γ) →
                cut p₀ close (wait (< q₀) P) ↝ P
\end{code}

@@

\begin{code}
  r-left      : ∀{Γ₁ Γ₂ A B P Q R}
                (p : Γ ≃ Γ₁ + Γ₂) (p₀ : Γ₁ ≃ [] + Γ₁) (q₀ : Γ₂ ≃ [] + Γ₂) →
                cut {A ⊕ B} p (left (< p₀) P)
                              (case (< q₀) Q R) ↝ cut p P Q
\end{code}

@@

\begin{code}[hide]
  r-right     : ∀{Γ₁ Γ₂ A B P Q R}
                (p : Γ ≃ Γ₁ + Γ₂) (p₀ : Γ₁ ≃ [] + Γ₁) (q₀ : Γ₂ ≃ [] + Γ₂) →
                cut {A ⊕ B} p (right (< p₀) P)
                              (case (< q₀) Q R) ↝ cut p P R
  r-fork      : ∀{Γ₁ Γ₂ Γ₃ Δ A B P Q R}
                (p : Γ ≃ Δ + Γ₃) (p₀ : Γ₃ ≃ [] + Γ₃) (q : Δ ≃ Γ₁ + Γ₂) (q₀ : Δ ≃ [] + Δ) →
                let _ , p′ , q′ = +-assoc-l p q in
                cut {A ⊗ B} p (fork (< q₀) q P Q)
                              (join (< p₀) R) ↝ cut q′ P (cut (> p′) Q R)
  r-client    : ∀{Γ₁ Γ₂ A P Q}
                (p : Γ ≃ Γ₁ + Γ₂) (p₀ : Γ₁ ≃ [] + Γ₁) (q₀ : Γ₂ ≃ [] + Γ₂) (un : Un Γ₁) →
                cut {`! A} p (server (< p₀) un P) (client (< q₀) Q) ↝ cut p P Q
  r-weaken    : ∀{Γ₁ Γ₂ A P Q}
                (p : Γ ≃ Γ₁ + Γ₂) (p₀ : Γ₁ ≃ [] + Γ₁) (q₀ : Γ₂ ≃ [] + Γ₂) (un : Un Γ₁) →
                cut {`! A} p (server (< p₀) un P)
                             (weaken (< q₀) Q) ↝ weakening un p Q
  r-contract  : ∀{Γ₁ Γ₂ A P Q}
                (p : Γ ≃ Γ₁ + Γ₂) (p₀ : Γ₁ ≃ [] + Γ₁) (q₀ : Γ₂ ≃ [] + Γ₂) (un : Un Γ₁) →
                cut {`! A} p (server (< p₀) un P)
                             (contract (< q₀) Q) ↝
                    contraction un p
                      (cut ++≃+ (server (< p₀) un P)
                      (cut (> p) (server (< p₀) un P) Q))
  r-poly       : ∀{A B Γ₁ Γ₂ P F} (p : Γ ≃ Γ₁ + Γ₂) (p₀ : Γ₁ ≃ [] + Γ₁) (q₀ : Γ₂ ≃ [] + Γ₂) ->
                 cut {`∃ A} p (ex {_} {B} (< p₀) P) (all (< q₀) F) ↝ cut p P (F B)
\end{code}

@@

\begin{code}
  r-cut        : ∀{Γ₁ Γ₂ A P Q R} (q : Γ ≃ Γ₁ + Γ₂) →
                 P ↝ Q → cut {A} q P R ↝ cut q Q R
  r-cong       : ∀{P R Q} → P ⊒ R → R ↝ Q → P ↝ Q
\end{code}
\end{AgdaAlign}
